### 2022-03-13 rmd/title.tex
edit: rmd

%\begin{titlepage}
\begin{center}
\LARGE{\textbf{Supplementary Materials for:}}\\[7.5mm]
\LARGE{Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks}\\[7.5mm]
\end{center}
\begin{flushleft}
\textbf{This PDF file includes:}
\begin{itemize}
\item XXX
\item References
\end{itemize}
\end{flushleft}
% \end{titlepage}

\newpage