### 2022-11-09 rmd/suppl_info.Rmd
edit

---
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
    includes:
      before_body: title.tex
mainfont: Times New Roman
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)

options(knitr.kable.NA = '')

# Sensitivity data
source(here::here("code/analysis_stvy.R"))
load(file = here::here("output/result_main.RData"))
load(file = here::here("output/result_stvy.RData"))

sim_sv <- sim_stvy_result
n_sim_stvy <- unique(sim_sv$n_warmup) + unique(sim_sv$n_burnin) + unique(sim_sv$n_timestep)

```

```{=tex}
\newpage
\pagenumbering{arabic}
```
# Supplementary text

## Sensitivity analysis

We performed a sensitivity analysis of the trophic community simulation to identify key simulation parameters that strongly affect the relationships between food chain length and ecosystem properties (the number of habitat patches $N_{p}$ and branching probability $P_{b}$. We generated 500 sets of parameter combinations by randomly drawing values of `r length(attributes(fit_sense$fit[[1]]$terms)$term.label)` simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we generated `r max(sim_sv$n_rep)` branching networks with the gradients of ecosystem size ($N_p \sim \text{Unif}(10, 150)$) and complexity ($P_b \sim \text{Unif}(0.01, 0.99)$). This results in a total of `r nrow(sim_sv)` simulation replicates. We ran `r n_sim_stvy` time steps of trophic community dynamics in each simulation replicate. We obtained temporal means of food chain length for the last `r unique(sim_sv$n_timestep)` time steps. The first `r n_sim_stvy - unique(sim_sv$n_timestep)` time steps were discarded as initialization and burn-in periods.

For each parameter combination, we estimated the Spearman's rank correlation between food chain length and ecosystem properties ($\psi_1$ for $N_{p}$ and $\psi_2$ for $P_{b}$). To examine influences of simulation parameters (**Table S1**), we developed the following regression model taking $\psi_1$ or $\psi_2$ as a response variable $u_n$ (parameter combination $n$):

```{=tex}
\begin{align}
  u_{n} &\sim \text{Normal}(\mu_n, \sigma^2)\\
  \mu_n &= \zeta_0 + \sum_{k}{\zeta_k x_{k, n}}
\end{align}
```
where $\zeta_k$ are the intercept and regression coefficients, and $x_{k,n}$ are the simulation parameters drawn from uniform distributions (**Table S1**). Response and explanatory variables were standardized to a mean of zero and a standard deviation of one so that regression coefficients are comparable.

The sensitivity analysis revealed key simulation parameters. For ecosystem size, dispersal distance $\theta$, local disturbance variation $\sigma_{l}$, and the reproductive rate of basal species $r_b$ were influential. However, parameters that influence food web structure were less influential (attack rate, conversion efficiency, and handling time). For ecosystem complexity, dispersal distance $\theta$ and probability $p_d$, local disturbance variation $\sigma_{l}$, and the reproductive rate of basal species $r_b$ were influential. Food web parameters were also important (attack rate, conversion efficiency, and handling time).

Based on the results, we considered ecological scenarios relevant for putative controls of food chain: low and high productivity ($r_b = 4, 8$), and low and high disturbance ($p_m = 0.01, 0.1$). We also investigated the role of omnivory by changing attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for predator $P$. Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$. To simplify simulation setup, we fixed the following parameters: $e_{BC}=$ `r unique(sim_main_result$e_bc)` , $a_{BC}=$ `r unique(sim_main_result$a_bc)`, and $h_{ij}=$ `r unique(sim_main_result$h_bc)`.

\pagebreak

# Tables

## Table S1 Parameter values

Parameter values used for the main and sensitivity analysis.

```{r table_param}

source(here::here("code/table_param.R"))
knitr::kable(df_value, format = "markdown")

```

\pagebreak

```{r table_sensitivity}

source(here::here("code/table_stvy.R"))

j <- 1

## conditional text
c_text <- c("ecosystem size",
            "ecosystem complexity")

table_stvy_chunk <- knit_chunk <- NULL

for(i in 1:length(table_stvy)) {

  table_legend <- paste0("Sensitivity analysis of the ", c_text[i]," effect on food chain length. Parameter estimates of linear regression models are shown. The response variable is the Spearman's rank correlation between food chain length and ", c_text[i],". Response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.\n")

  knit_chunk <- paste0("## Table S", i + j, " Sensitivity analysis for ", c_text[i],
                       "\n\n",
                       table_legend,
                       "```{r table_s", i + j,", results='asis'}
                       \n\n
                       table_stvy[[", i,"]]
                       \n\n",
                       "```",
                       "\n\\pagebreak\n")

  table_stvy_chunk <- c(table_stvy_chunk, knit_chunk)

}

```

`r paste(knit(text = table_stvy_chunk), collapse = '\n')`

\pagebreak

# Figures

```{r si_figure_fcl}

source(here::here("code/figure_si_fcl.R"))

j <- 0

## conditional text
c_text <- c("ecosystem size",
            "ecosystem complexity")

fig_fcl_chunk <- knit_chunk <- NULL

for(i in 1:length(list_g_fcl)) {

  K <- length(list_g_fcl[[i]])

  for(k in 1:K) {

      figure_legend <- paste0(str_to_sentence(c_text[i])," effect on food chain length and occupancies of constituent species (proportion of habitat patches occupied). Rows are different response variables, and columns are different simulation scenarios. Lines and shades represent loess curves fitted to the simulated data and associated 95% confidence intervals. In this simulation, the following values of parameters were used: $\\theta$ = ", df_param$theta[k], ", $\\sigma_l$ = ", df_param$sd_disturb_lon[k], ", and $\\sigma_h$ = ", df_param$sd_disturb_source[k], ".")

  knit_chunk <- paste0("## Figure S", K * (i - 1) + k + j, " ", str_to_sentence(c_text[i]), " effect on food chain length",
                       "\n\n",
                       "```{r figure_s", K * (i - 1) + k + j,", fig.height = 10, fig.width = 11.5}",
                       "\n\n
                       list_g_fcl[[", i, "]][[", k, "]]
                       \n\n",
                       "```",
                       "\n\n",
                       figure_legend,
                       "\n\\pagebreak\n")

  fig_fcl_chunk <- c(fig_fcl_chunk, knit_chunk)

  }

}

```

`r paste(knit(text = fig_fcl_chunk), collapse = '\n')`