### 2023-03-08 rmd/suppl_info.Rmd
update: si document

---
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
    includes:
      before_body: title.tex
mainfont: Times New Roman
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
fontsize: 11pt
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)

options(knitr.kable.NA = '')

df_sim1 <- readRDS(here::here("output/sim_one_patch.rds"))
n_sim <- df_sim1 %>% 
  mutate(x = n_timestep + n_warmup + n_burnin,
         y = n_warmup,
         z = n_burnin) %>% 
  dplyr::select(x, y, z) %>% 
  distinct() %>% 
  unlist()
```

```{=tex}
\newpage
\pagenumbering{arabic}
```
# Supplementary text

## One-patch simulation

We performed one-patch simulations to identify parameter combinations that allow stable coexistence of the three species without dispersal and disturbance. Specifically, we varied attack rates $a_{ij}$, handling time $h_{ij}$, intrinsic growth rate of the basal species $r_b$, and carrying capacity $K$. First, we considered `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` discrete combinations of $a_{BC}$, $r_b$, $K$, and $h$ (assumed $h_{ij} = h$). We further crossed `r nrow(dplyr::distinct(df_sim1, a2))` values of $a_{BP}$ (`r min(df_sim1$a2)` -- `r max(df_sim1$a2)` with an equal interval) and $a_{CP}$ (`r min(df_sim1$a3)` -- `r max(df_sim1$a3)`). Consequently, we had a total of `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, a2, a3, h))` combinations for one-patch simulations.

For each parameter combination, we ran 10 replications of `r n_sim[1]` time steps, which include `r n_sim[2]` warmup and `r n_sim[3]` burn-in periods. We recorded the patch state as numeric values ("No species" = 0, "B" = 1, "B + C" = 2, "B + P" =3, and "B + C + P" = 4) at $t = 1600$ and retrieved the maximum value of 10 replicates.

Results were summarized in **Figures S1 and S2**. We identified the following three parameter combinations that could maintain the state of "B + C + P" with distinct food web motifs: "chain (no omnivory)" $(a_{BP}, a_{CP}, h) = (0, 0.0025, 0.5)$, "weak omnivory" ($0.02, 0.025, 0.5$), and "strong omnivory" ($0.04, 0.0025, 0.75$) (see symbols in **Figures S1 and S2**). In the maintext, we reported the results with $a_{BC} = 0.5$ because qualitative patterns did not depend on $a_{BC}$.

\newpage

# Figures

```{r fig1, fig.cap="Results of one-patch simulation with $a_{BC} = 0.25$. Colors indicate the patch state, and points with different symbols identify parameter combinations used in the main simulation. Panels distinguishh the value of carrying capacity $K$. In both panels, rows represent different values of intrinsic growth rate $r_b$, and columns identify handling time $h$."}
knitr::include_graphics(here::here("figure/figure_si_one_patch_a1_25.pdf"))
```

\newpage

```{r fig2, fig.cap="Results of one-patch simulation with $a_{BC} = 0.50$. Colors indicate the patch state, and points with different symbols identify parameter combinations used in the main simulation. Panels distinguishh the value of carrying capacity $K$. In both panels, rows represent different values of intrinsic growth rate $r_b$, and columns identify handling time $h$."}
knitr::include_graphics(here::here("figure/figure_si_one_patch_a1_50.pdf"))
```

\newpage

```{r fig3, fig.cap="Relationships between food chain length and ecosystem size with low disturbance intensity $\\mu_m = 0.2$. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios."}
knitr::include_graphics(here::here("figure/figure_si_n_patch.pdf"))
```

\newpage

```{r fig4, fig.cap="Relationships between food chain length and ecosystem complexity with low disturbance intensity $\\mu_m = 0.2$. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios."}
knitr::include_graphics(here::here("figure/figure_si_p_branch.pdf"))
```

\newpage

```{r fig5, fig.cap="Relationships between patch state occupancy and ecosystem size with high disturbance intensity $\\mu_m = 0.8$. Panels indicate different food web motifs. In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios."}
knitr::include_graphics(here::here("figure/figure_si_state_n_patch.pdf"))
```

\newpage

```{r fig6, fig.cap="Relationships between patch state occupancy and ecosystem complexity with high disturbance intensity $\\mu_m = 0.8$. Panels indicate different food web motifs. In each panel, rows distinguish patch state, and columns indicate different dispersal and productivity scenarios."}
knitr::include_graphics(here::here("figure/figure_si_state_p_branch.pdf"))
```

\newpage

```{r fig7, fig.cap="Effects of productivity (left column) and dispersal (right column) on food chain length grouped by disturbance frequency ($p_m$; x-axis). Colors distinguish productivity or dispersal scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
knitr::include_graphics(here::here("figure/figure_si_fcl_r_disp.pdf"))
```

\newpage

```{r fig8, fig.cap="Effects of productivity on state transitions. Colors distinguish productivity scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
knitr::include_graphics(here::here("figure/figure_si_state_r.pdf"))
```

\newpage

```{r fig9, fig.cap="Effects of dispersal on state transitions. Colors distinguish distinguish scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
knitr::include_graphics(here::here("figure/figure_si_state_disp.pdf"))
```