### 2023-03-08 rmd/title.tex
update: si

%\begin{titlepage}
\begin{center}
\LARGE{\textbf{Supplementary Materials for:}}\\[7.5mm]
\LARGE{Ecosystem size and complexity as extrinsic drivers of food chain length in branching networks}\\[7.5mm]
\large{Justin P.F. Pomeranz, Jacques C. Finlay, Akira Terui}\\[7.5mm]
\end{center}
\begin{flushleft}
\textbf{This PDF file includes:}
\begin{itemize}
\item Supplementary text
\item Tables S1 - S3
\item Figures S1 - S4
\end{itemize}
\end{flushleft}
% \end{titlepage}

\newpage