### 2023-05-17 rmd/suppl_info.Rmd
edit: text

---
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
    includes:
      before_body: title.tex
mainfont: Times New Roman
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
fontsize: 11pt
header-includes:
  - \usepackage{amsmath, fancyhdr, caption, booktabs}
  - \pagestyle{fancy}
  - \renewcommand{\headrulewidth}{0pt}
  - \pagenumbering{gobble}
  - \setlength{\headheight}{14pt}
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)

options(knitr.kable.NA = '')

df_sim1 <- readRDS(here::here("output/sim_one_patch.rds"))
n_sim <- df_sim1 %>% 
  mutate(x = n_timestep + n_warmup + n_burnin,
         y = n_warmup,
         z = n_burnin) %>% 
  dplyr::select(x, y, z) %>% 
  distinct() %>% 
  unlist()
```

```{=tex}
\newpage
\pagestyle{fancy}
\fancyhead[L]{Pomeranz et al.}
\fancyhead[C]{}
\renewcommand{\headrulewidth}{0.5pt}
\pagenumbering{arabic}
\def\figurename{Figure S\hspace{-.4ex}}
\def\tablename{Table S\hspace{-.4ex}}
```
# Supplementary text

## One-patch simulation

We performed one-patch simulations to identify parameter combinations that allow stable coexistence of the three species without dispersal and disturbance. Specifically, we varied attack rates $a_{ij}$, handling time $h_{ij}$, intrinsic growth rate of the basal species $r_b$, and carrying capacity $K$. First, we considered `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` discrete combinations of $a_{BC}$, $r_b$, $K$, and $h$ (assumed $h_{ij} = h$). We further crossed `r nrow(dplyr::distinct(df_sim1, a2))` values of $a_{BP}$ (`r min(df_sim1$a2)` -- `r max(df_sim1$a2)` with an equal interval) and $a_{CP}$ (`r min(df_sim1$a3)` -- `r max(df_sim1$a3)`), resulting in `r nrow(dplyr::distinct(df_sim1, a2, a3))` combinations. Consequently, we had a total of `r nrow(dplyr::distinct(df_sim1, r_b, k, a1, a2, a3, h))` combinations (`r nrow(dplyr::distinct(df_sim1, r_b, k, a1, h))` $\times$ `r nrow(dplyr::distinct(df_sim1, a2, a3))`) for one-patch simulations. Note that we set the switching parameter $s = 0$ if $a_{BP} = 0$; otherwise $s = 1$. Parameter values were summarized in **Table S1**.

For each parameter combination, we ran ten replications of `r n_sim[1]` time steps, which include `r n_sim[2]` warmup and `r n_sim[3]` burn-in periods. We recorded the patch state as numeric values ("No species" = 0, "B" = 1, "B + C" = 2, "B + P" =3, and "B + C + P" = 4) at $t = 1600$ and retrieved the maximum value of ten replicates.

Results were summarized in **Figures S1 and S2**. We identified the following three parameter combinations that could maintain the state of "B + C + P" with distinct food web motifs: "chain (no omnivory)" $(a_{BP}, a_{CP}, h) = (0, 0.0025, 0.5)$, "weak omnivory" ($0.02, 0.025, 0.5$), and "strong omnivory" ($0.04, 0.0025, 0.75$) (see symbols in **Figures S1 and S2**). In the main text, we reported the results with $a_{BC} = 0.5$ because qualitative patterns did not depend on $a_{BC}$.

\newpage

# Tables

```{r}
source(here::here("code/table_param.R"))

kable(df_value,
      escape = FALSE,
      booktabs = TRUE,
      caption = "Parameter values used in the main and one-patch simulation") %>% 
  kableExtra::kable_styling(latex_options = 'HOLD_position')
```

\newpage

# Figures

```{r fig1, fig.cap="Results of one-patch simulation with $a_{BC} = 0.25$. Colors indicate the patch state, and points with different symbols identify parameter combinations used in the main simulation. Panels distinguish the value of carrying capacity $K$. In both panels, rows represent different values of intrinsic growth rate $r_b$, and columns identify handling time $h$."}
knitr::include_graphics(here::here("figure/figure_si_one_patch_a1_25.pdf"))
```

\newpage

```{r fig2, fig.cap="Results of one-patch simulation with $a_{BC} = 0.50$. Colors indicate the patch state, and points with different symbols identify parameter combinations used in the main simulation. Panels distinguish the value of carrying capacity $K$. In both panels, rows represent different values of intrinsic growth rate $r_b$, and columns identify handling time $h$."}
knitr::include_graphics(here::here("figure/figure_si_one_patch_a1_50.pdf"))
```

\newpage

```{r fig3, fig.cap="Relationships between food chain length and ecosystem size with low disturbance intensity $\\mu_m = 0.2$.  Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities. In this figure, ecosystem complexity values for individual simulation replicates were drawn randomly from a uniform distribution $P_b \\sim \\text{Unif}(0.01, 0.99)$."}
knitr::include_graphics(here::here("figure/figure_si_n_patch.pdf"))
```

\newpage

```{r fig4, fig.cap="Relationships between food chain length and ecosystem complexity with low disturbance intensity $\\mu_m = 0.2$. Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities. In this figure, ecosystem complexity values for individual simulation replicates were drawn randomly from a uniform distribution $N_p \\sim \\text{Unif}(10, 150)$."}
knitr::include_graphics(here::here("figure/figure_si_p_branch.pdf"))
```

\newpage

```{r fig5, fig.cap="Effects of productivity (left column) and dispersal (right column) on food chain length grouped by disturbance frequency ($p_m$; x-axis). Box plots show median (horizontal center line) and quartiles (box limits) of $500$ simulation replicates with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range. Dots are outliers extending beyond ±1.5 interquartile range. Colors distinguish productivity or dispersal scenarios, and rows identify food web motifs. Disturbance intensity $\\mu_m = 0.8$."}
knitr::include_graphics(here::here("figure/figure_si_fcl_r_disp.pdf"))
```

\newpage

```{r fig6, fig.cap="Effects of productivity on state transitions with high disturbance intensity $\\mu_m = 0.8$. The proportion of each patch state $\\bar{\\psi}$ is averaged over time for each simulation replicate. Box plots show median (horizontal center line) and quartiles (box limits) of $500$ simulation replicates with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range. Dots are outliers extending beyond ±1.5 interquartile range. Colors distinguish productivity scenarios, columns represent different disturbance probability $p_m$, and rows identify food web motifs."}
knitr::include_graphics(here::here("figure/figure_si_state_r.pdf"))
```

\newpage

```{r fig7, fig.cap="Effects of dispersal on state transitions with high disturbance intensity $\\mu_m = 0.8$. The proportion of each patch state $\\bar{\\psi}$ is averaged over time for each simulation replicate. Box plots show median (horizontal center line) and quartiles (box limits) of $500$ simulation replicates with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range. Dots are outliers extending beyond ±1.5 interquartile range. Colors distinguish dispersal scenarios, columns represent different disturbance probability $p_m$, and rows identify food web motifs."}
knitr::include_graphics(here::here("figure/figure_si_state_disp.pdf"))
```

\newpage

```{r fig8, fig.cap="Ecosystem size and complexity dictate total carrying capacity (= $\\sum_x K_x$) and mean distance between a given pair of habitat patches within a network."}
knitr::include_graphics(here::here("figure/figure_si_k_distance.pdf"))
```