### 2023-07-10 rmd/readme.Rmd
edit: readme

---
title: "README"
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here())
      })
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)

pacman::p_load(tidyverse)
```

## Article Information

Title: Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks

## File descriptions

`code/` includes all R codes to reproduce findings from this research:

```{r}

list.files(here::here("code")) %>% 
  as_tibble() %>% 
  mutate(Description = case_when(str_detect(value, "^figure_")~"Code for figures: '_concept', Figure 1; '_delta', Figure 1D; '_food_web', Figure 1C; '_treeplot', Figure2; '_main_geom', Figure 3, 4; '_network', Figure 1A; '_patch_state', Figure 5-10",
                                 str_detect(value, "^figure_si")~"Code for SI figures: '_geom', Figure S3, S4; '_one_patch', Figure S1, S2; '_r_disp', Figure S5; '_within_net', Figure S8",
                                 str_detect(value, "^format_")~"Code for formatting",
                                 str_detect(value, "^library")~"Code for figures",
                                 str_detect(value, "^run_")~"Code for simulations",
                                 str_detect(value, "^table_")~"Code for tables",
                                 str_detect(value, "^set_")~"Code for setting figure themes"
                                 )) %>% 
  rename(Name = value) %>% 
  knitr::kable(format = "markdown")
```

## Session Information

```{r}

source(here::here("code/library.R"))
sessionInfo()

```